---
title: "Dataset Selection"
author: "Raghav, Malik, Himansu, Ari"
date: "10/24/2021"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Business Problem

### What business problem are you addressing? Remember - we take a very wide and broad view of what is "business". You should describe the context - who is your audience, what decision do they need to make, why is this problem significant/important?
    
The business problem that we hope to address with our project is the cost of losing customers. Our target audience is the management of telecom companies, as the findings of our analysis will help them identify which customers are most likely to cease use of their service ("churn"). If management can successfully identify these individuals, then they can construct targeted promotions and advertising campaigns focused on this precarious segment. This problem is significant because the telecom industry is extremely competitive and usually relies on customer retention for revenue and long-term profitability. When a customer is lost, this harms the company not only because they lose a source of income, but also since many infrastructure costs remain constant despite one fewer customer using the service. 
    

    

## Dataset
```{r}
data <- read.csv("WA_Fn-UseC_-Telco-Customer-Churn.csv", stringsAsFactors = TRUE)
summary(data)
```


### What dataset are you using? Include details - source of the data, size of the data, what is your response value, what are the predictors?

Source: Kaggle
    
Size: 7043 rows (customers) and 21 columns (features)
    
Response: "Churn" column
    
Predictors:

<ul>
  <li><b>Information regarding services</b> that each customer has signed up for (phone, multiple lines, internet, online security, online backup, device protection, tech support, and streaming TV and movies)</li>
  <li><b>Customer account information</b> (how long they’ve been a customer, contract, payment method, paperless billing, monthly charges, and total charges)</li>
  <li><b>Demographic info</b> about customers (gender, age range, and if they have partners and dependents)</li>
</ul>

## Challenges

### Any challenges that are clear at this point - data cleaning issues, NA issues, data size issues, what else?
    
Some irrelevant data such as customer ID should be omitted from the dataset as it has no bearing on whether a customer will churn or not. In addition, we notice from performing a quick read-and-summary that the column "TotalCharges" has 11 N/A values, which should likely not exist since all customers have data for monthly charges. This data could be replaced with a value equal to tenure*monthlyCharge, or perhaps it would be better to completely eliminate those rows due to the fact that there are very few rows with missing data out of the total 7043 rows.


## Why this project

### Lastly - but most importantly, why should the client spend money and resources to do this project?

By predicting which customers are most likely to leave, companies can target them with an offer or promotion that incentivizes them to continue using their service. Although offering a promotion cuts into profit, if structured properly, the promotional expenditure will be small relative to the incremental contraction in revenue associated with the loss of a customer. Therefore, by accurately predicting which customers are most likely to leave and successfully offering them incentives to remain on the platform, telecom companies can significantly increase their profits. 

It should be noted that there is risk associated with this strategy, particularly in the case of a false positive. If a company identifies a customer as likely to churn when they actually are not, offering a promotion to such a customer would result in the company losing profit. Understanding the danger of false positives will help guide our analysis, and may be reason for us to select a more conservative model as opposed to one which identifies potential churn candidates more liberally. We look forward to meticulously analyzing the data at hand and performing an analysis that adds significant value to our client's business model.
